### 2019-03-04 README.Rmd
analysis -> public

---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Estimating the effect of the 2005 change in BCG policy in England: A retrospective cohort study

**Background**

In 2005, England changed from universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children to targeted BCG vaccination of high-risk children at birth.

**Methods**

We combined notification data from the Enhanced Tuberculosis Surveillance system, with demographic data from the Labour Force Survey to construct retrospective cohorts of individuals in England relevant to both the universal, and targeted vaccination programmes between Jan 1, 2000 and Dec 31, 2010. For each cohort, we estimated incidence rates over a 5 year follow-up period and used Poisson and Negative Binomial regression models in order to estimate the impact of the change in policy on TB.

**Results:**

In the non-UK born, we found evidence for an association between a reduction in incidence rates and the change in BCG policy (school-age IRR: 0.74 (95%CI 0.61, 0.88), neonatal IRR: 0.62 (95%CI 0.44, 0.88)). We found some evidence that the change in BCG policy was associated with a increase in incidence rates in the UK born school-age population (IRR: 1.08 (95%CI 0.97, 1.19)) and weaker evidence of an association with a reduction in incidence rates in UK born neonates (IRR: 0.96 (95%CI 0.82, 1.14)). Overall, we found that the change in BCG policy was associated with directly preventing 385 (95% CI -105, 881) TB cases.

**Conclusions**

The change in BCG policy was associated with reduced incidence of TB in the study population. This was largely driven by reductions in the non-UK born. There was a slight increase in those at school-age, who were UK born.

# Docker

This analysis was developed in a docker container based on the tidyverse docker image. To run the docker image run:

```{bash, eval = FALSE}
docker run -d -p 8787:8787 --name DirectEffBCGPolicyChange -e USER=DirectEffBCGPolicyChange -e PASSWORD=DirectEffBCGPolicyChange seabbs/directeffbcgpolicychange
```

The rstudio client can be found on port :8787 at your local machines ip. The default username:password is DirectEffBCGPolicyChange:DirectEffBCGPolicyChange, set the user with -e USER=username, and the password with - e PASSWORD=newpasswordhere. The default is to save the analysis files into the user directory.

If you have access to the required underlying raw data then the entire analysis can be reproduced from scratch by adding the following to the `docker run` command, with the data saved into `data/tb_data`. The data requirements, and structure, can be found [here](https://www.samabbott.co.uk/tbinenglanddataclean/).

```{bash, eval = FALSE}
--mount type=bind,source=$(pwd)/data/tb_data,target=/home/DirectEffBCGPolicyChange/data/tb_data
```

### 2019-03-04 R/Base_Model.R

@@ -1,0 +1,55 @@
+## Model function to be used across results
+## Set MCMC options across models as function defaults
+#' Generic BRM model with fitting defaults set
+#'
+#' @description A generic model used to define default \code{brms} settings to be used across all
+#' models investigated.
+#' @param Predictors A character vector of variables to use as features
+#' @param ModelFamily A character string containing the model family to fit.
+#' @param Data A dataframe of the data to fit to.
+#' @param SetIter Numeric, define the number of iterations
+#' @param SetBurn Numeric, define the number of burn in iterations
+#' @param SetCores Numeric, set the number of cores to use.
+#' @param Chains Numeric, set the number of chains to use for MCMC
+#' @param SetThin Nuemeric, set the number of iterations to thin over.
+#' @param Silent Logical, defaults to \code{TRUE}. Should the results be returned without progress messages.
+#' @param SaveModel Logical, defaults to \code{TRUE}. Should the model output be saved
+#' @param AdaptDelta Numeric, adaptive delta to use
+#' @param MaxTree Numeric, the maximum number of trees.
+#'
+#' @return A \code{brm} fitted model.
+#' @export
+#' @seealso ModelConstruct
+#' @importFrom brms brm
+#' @import rstan
+#' @import Rcpp
+#' @importFrom parallel detectCores
+#' @examples
+#' 
+Base_Model = function(Predictors = NULL, ModelFamily, Data = NULL, SetIter = NULL, SetBurn = NULL,
+                      SetCores = NULL, Chains = 4, SetThin = 1,  Silent = TRUE, SaveModel = TRUE,
+                      AdaptDelta  = NULL, MaxTree = NULL) {
+  
+  # Load Rstan
+  # set no predictor option
+  if (is.null(Predictors)) {
+    Predictors <- '1'
+  }
+  
+  if (is.null(SetCores)) {
+    SetCores <- parallel::detectCores() - 1 
+  }
+  ## Stop if no data
+  if (is.null(Data)) {
+    stop('No data has been specified')
+  }
+  #Specify model
+  Model <- Predictors %>% ModelConstruct
+  ## Fit model
+  ModelFit <- brm(as.formula(Model), family = ModelFamily, data = Data, chains = Chains,
+                  cores = SetCores, iter = SetIter, warmup = SetBurn, thin = SetThin, 
+                  refresh = ifelse(Silent, 0, 1), save_dso = SaveModel, 
[…]

### 2019-03-04 R/ModelConstruct.R

@@ -1,0 +1,12 @@
+#' Specify the Model Formula using Population as an Offset and Cases as the Outcome
+#'
+#' @param Preds A character vector containing variable names.
+#'
+#' @return A character string containing the model formula
+#' @export
+#'
+#' @examples
+#' 
+ModelConstruct <- function(Preds) {
+  paste0('Cases ~ ', paste(Preds, collapse = ' + '), ' + offset(log(Population))')
+}

### 2019-03-04 R/OR_inf_rep.R

@@ -1,0 +1,30 @@
+#' Inf Replacement for Large Odds Ratios
+#'
+#' @param OR Numeric, or character odds ratio
+#' @param MaxSize Numeric, maximum size for odds ratio. Defaults to 
+#' 100000
+#'
+#' @return A bounded effect estimate or Inf.
+#' @export
+#'
+#' @examples
+#' 
+OR_inf_rep = function(OR, MaxSize = 100000) {
+  OR <- as.numeric(OR)
+  OR <- ifelse(OR > MaxSize, Inf, OR)
+  return(OR)
+}
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2019-03-04 R/UnloadUserDLLs.R

@@ -1,0 +1,27 @@
+#' Unload Named DLLs
+#'
+#' @param VectorDLLNames A character vetor of DLL names.
+#' @param verbose Logical defautls to \code{TRUE}. Should progress messages be printed.
+#'
+#' @return This functions returns \code{NULL}
+#' @export
+#' @importFrom purrr map
+#' @examples
+#' 
+UnloadUserDLLs = function(VectorDLLNames, verbose = TRUE) {
+  
+  DLLSummary <- getLoadedDLLs()
+  
+  VectorDLLNames %>% 
+    map(~DLLSummary[[.]][['path']]) %>%
+    map(dyn.unload)
+  
+  if (verbose & length(VectorDLLNames) != 0) {
+    
+    message('A total of ', VectorDLLNames %>% 
+              length, ' DLL have been unloaded. These were ',
+            paste(VectorDLLNames, collapse = ', ') ,
+            '. This may restrict the future use of linked objects.')
+  }
+  
+}

### 2019-03-04 R/add_looic_com_to_no_policy_change.R

@@ -1,0 +1,34 @@
+#' Add LOOIC Comparision for a Model Without the Effect of the Change in Policy
+#'
+#' @param model A model object as produced by \code{best_fit_model}
+#' @param reloo Logical, defaults to \code{TRUE}. Should problematic 
+#' observations be refit.
+#' @param cores Numeric, defaults to 1. The number of cores to use.
+#'
+#' @return A list containing the fitted model object and a tibble containing the 
+#' comparision LOOIC scores and standard error
+#' @export
+#' @importFrom tibble tibble
+#' @importFrom brms LOO
+#' @importFrom prettypublisher pretty_round
+#' @examples
+#' 
+add_looic_com_to_no_policy_change <- function(model, reloo = TRUE, cores = 1) {
+  
+  model_no_policy_change <- update(model$Fit, . ~ . - PolicyChange, cores = cores)
+  
+  loo_ic <- LOO(model$Fit, model_no_policy_change, reloo = reloo, cores = cores)
+  
+  out <- tibble(diff_looic = loo_ic$ic_diffs__[1],
+                diff_looic_se = loo_ic$ic_diffs__[2]) %>% 
+    mutate(
+      pretty_diff_looic = paste0(pretty_round(diff_looic, 2), 
+                                 " (SE ",
+                                 pretty_round(diff_looic_se, 2),
+                                 ")")
+    )
+  
+  model$diff_drop_policy_change <- out
+   
+  return(model)            
+}

### 2019-03-04 R/best_fit_results_table.R

@@ -1,0 +1,57 @@
+#' Function to generate results table
+#'
+#' @param df A dataframe of model results
+#' @param Scheme A character string specifying the vaccination scheme to make the 
+#' table for.
+#' @param Pop A character string indicating which population to make the results table for.
+#' @param model_to_select Numeric, defaults to 1. Manually select which model to extract. Useful if several models
+#' are nearly comparable.
+#' @seealso model_comparision
+#' @return A list containing the best fitting model name, the model caption, the tidied effect size 
+#' output and the summarised model for inspection
+#' @export
+#'
+#' @inheritParams format_model_summaries
+#' @importFrom dplyr first filter select mutate rowwise rename
+#' @importFrom purrr flatten
+#' @importFrom tibble as_data_frame
+#' @importFrom stringr str_replace
+#' @examples
+#' 
+best_fit_results_table = function(df, Scheme, Pop,
+                                  sep = ", ",
+                                  model_to_select = 1) {
+  ## Run model comparision
+  ModelCom <- model_comparison(df, Scheme, Pop)
+  
+  ## Best fitting model
+  BestFitModel <- ModelCom %>% first %>% nth(model_to_select)
+  
+  ## Filter data for this model
+  df <- df %>% 
+    filter(VacScheme %in% Scheme, CoB %in% Pop, Model %in% BestFitModel)
+  
+  ## Save caption
+  Caption <- df %>% select(Description) %>% unlist
+
+  ## Save Tidied effect size output
+  TidySum <- df %>% 
+    select(TidySum) %>%
+    flatten %>%
+    flatten %>% 
+    as_data_frame %>% 
+    mutate(term = str_replace(term, '_', ' ')) %>% 
+    mutate(term = str_replace(term, 'b ', '')) %>% 
+    mutate(term = str_replace(term, 'r ', '')) %>%
+    mutate(term = str_replace(term, '__', ' ')) %>%
+    rowwise %>% 
+    mutate(estimate = c(estimate, lower, upper) %>% 
+             as.character %>%
+             pretty_ci(string = TRUE, sep = sep)) %>%
+    select(-std.error, -lower, -upper)
+
+  
+  return(list(BestFitModel, Caption, TidySum))   
+}
+
+

### 2019-03-04 R/best_model_fit.R

@@ -1,0 +1,36 @@
+#' Return Model Fit for Plotting
+#'
+#' @description Extracts a given model and refits it, returning the fitted model object.
+#' @param df A dataframe of fitted model results
+#' @param ModelBestFit A character string identifying the model to refit.
+#' @param Scheme A character string identifying which scheme to refit the model for.
+#' @param Pop A character string identifying the population to refit the model for.
+#' @inheritParams fit_stan_models
+#' @seealso Base_Model
+#' @return Fitted model object
+#' @export
+#' @importFrom dplyr filter do select
+#' @importFrom purrr flatten
+#' @examples
+#' 
+best_model_fit = function(df = NULL, ModelBestFit = NULL, 
+                          Scheme = NULL, Pop = NULL, Cores = NULL, 
+                          Iterations = 10000, BurnIn = 5000, 
+                          Alpha = 0.99, Treedepth = 10) {
+  ## Filter data for this model
+  ModelFit <- df %>% 
+    filter(VacScheme %in% Scheme, CoB %in% Pop, Model %in% ModelBestFit) %>% 
+    do(
+      
+      Fit = Base_Model(.$Predictors %>% first,
+                       ModelFamily = .$ModelFamily %>% first, 
+                       Data = .,
+                       Silent = TRUE,  SetIter = Iterations, SetBurn =BurnIn,
+                       SetCores = Cores, Chains = 4,
+                       AdaptDelta  = Alpha, MaxTree = Treedepth
+      )
+    ) %>% 
+    select(Fit) %>% 
+    flatten 
+  return(ModelFit)
+}

### 2019-03-04 R/capture_stan_DLLs.R

@@ -1,0 +1,33 @@
+#' Function to print the number of stan models loaded as DLLs and return their names
+#'
+#' @param verbose Logical, defaults to \code{TRUE}. If \code{TRUE} the number of loaded stan
+#' models are printed.
+#' @param StanFileNames A character string identifying the Stan file names, defaults to file.
+#' @param MaxDLLs Numeric, DLL cap. The maximum defaults to 100.
+#'
+#' @return A character vector of the currently loaded DLLs.
+#' @export
+#' @importFrom purrr map_lgl
+#' @examples
+#' 
+capture_stan_DLLs = function(verbose = TRUE, StanFileNames = 'file', MaxDLLs = 100) {
+  
+  ## Capture DLLs
+  DLLs <- getLoadedDLLs()
+  
+  StanFile <- grepl(StanFileNames, DLLs)
+  
+  ## Summmary of DLL breakdown
+  TotalStanDLLs <- StanFile %>% sum
+  
+  TotalNonStanDLLs <- StanFile %>% map_lgl(~!.) %>% sum
+  
+  ## Message summarising DLLs
+  if (verbose) {
+    message('There are currently ', TotalStanDLLs, ' stan models loaded as DLLs, with ', TotalNonStanDLLs, ' DLLs loaded that are not stan models. There are ', MaxDLLs - TotalStanDLLs - TotalNonStanDLLs, ' DLLs available before the DLL cap is reached')
+  }
+  
+  StanFileNames <- names(DLLs)[StanFile]
+  
+  return(StanFileNames)
+}

### 2019-03-04 R/case_rate.R

@@ -1,0 +1,72 @@
+#' Function to estimate the case rate for a given variable
+#'
+#' @description Estimates the case rate for a given variable, stratified by given variables,
+#'  dependant on age, year, agegrp2, and uk birth status. Missing data for the rate variable
+#'   is dropped.
+#'
+#' @param df A dataframe
+#' @param rate_for A character string indicating the variables to estiamte the rate for.
+#' @param strat_by A character string indicating the variables to stratify by.
+#' @param age_split A character string, defaults to \code{NULL}. Specifies the age group to
+#' stratify for.
+#' @param CoB_split Logical, defaults to \code{TRUE}. Should the case rate be grouped by the country of
+#' birth.
+#' @param Year_strat Logical, defaults to \code{TRUE}. Should the case rate be grouped by year.
+#'
+#' @seealso extract_case_counts
+#' @return A dataframe of case rates with upper and lower bounds
+#' @export
+#'
+#'
+#' @importFrom dplyr filter_at group_by summarise ungroup group_by rowwise mutate
+#' @examples
+#' 
+case_rate = function(df, rate_for, strat_by = NULL, age_split = NULL,
+                     CoB_split = TRUE, Year_strat = TRUE) {
+  CaseCountVar <- c('age', 'year', 'ukborn')
+  GroupByCase <- c()
+  GroupByTotCase <- c()
+
+  if (!is.null(strat_by)) {
+
+    CaseCountVar <- c(CaseCountVar, strat_by)
+    GroupByCase <- c(GroupByCase, strat_by)
+    GroupByTotCase <- c(GroupByTotCase, strat_by)
+  }
+
+  if (!is.null(rate_for)) {
+
+    CaseCountVar <- c(CaseCountVar, rate_for)
+    GroupByCase <- c(GroupByCase, rate_for)
+  }
+
+  if (!is.null(age_split)) {
+    GroupByCase <- c(GroupByCase, age_split)
+    GroupByTotCase <- c(GroupByTotCase, age_split)
+  }
+
+  if (CoB_split) {
+    GroupByCase <- c(GroupByCase, 'CoB')
+    GroupByTotCase <- c(GroupByTotCase, 'CoB')
+  }
+
+  if (Year_strat) {
+    GroupByCase <- c(GroupByCase, 'Year')
+    GroupByTotCase <- c(GroupByTotCase, 'Year')
+  }
+
+
[…]

### 2019-03-04 R/clear_user_DLL.R

@@ -1,0 +1,18 @@
+#' Clear Stan DLLs
+#'
+#' @description A wrapper function around \code{capture_stan_DLLS} and \code{UnloadUserDLLs}.
+#' Identifies the loaded Stan DLLs and unloads them.
+#' @inheritParams capture_stan_DLLs 
+#'
+#' @seealso capture_stan_DLLs UnloadUserDLLs
+#' @return This functions returns \code{NULL}
+#' @export
+#'
+#' @examples
+#' 
+clear_user_DLL = function(verbose = TRUE) {
+  
+  tmp <- capture_stan_DLLs(verbose)
+  
+  UnloadUserDLLs(tmp, verbose)
+}

### 2019-03-04 R/epi_conf.R

@@ -1,0 +1,17 @@
+#' Estimate Incidence Rates per 100,000
+#'
+#' @param Cases Numeric vector of cases
+#' @param Population Numeric vector of populations
+#'
+#' @return Estimated incidence rates with confidence intervals.
+#' @export
+#'
+#' @importFrom epiR epi.conf
+#' @examples
+#' 
+epi_conf = function(Cases, Population) {
+  epi.conf(matrix(c(sum(Cases, na.rm = TRUE), 
+                    sum(Population, na.rm = TRUE)), 
+                  ncol = 2, nrow = 1), ctype = "inc.rate",
+           method = "exact", N = 100000)
+}

### 2019-03-04 R/extract_inf_ef.R

@@ -1,0 +1,19 @@
+#' Extract incidence estimates
+#'
+#' @param df A dataframe containing a Model and IRR variables.
+#' @param model A Character string identifying the model to extract
+#'
+#' @return Returns a vector of formatted IRRs formatted for publication
+#' @export
+#'
+#'
+#' @importFrom prettypublisher pretty_inline_ci
+#' @importFrom dplyr filter select
+#' @examples
+#' 
+extract_inc_ef = function(df, model) {
+  df %>% 
+    filter(Model %in% model) %>% 
+    select(`IRR (CI 95%)*`) %>%
+    unlist %>% pretty_inline_ci(note = "95%CI ")
+}

### 2019-03-04 R/fit_stan_models.R

@@ -1,0 +1,256 @@
+#' Fit and summarise models
+#'
+#' @param ModelReadyDataSpec A dataframe of data ready for modelling
+#' @param PreCompile Logical, defaults to \code{TRUE}. Should a subset of models be precompiled
+#' (specified with \code{BaseLineModelNames}) for the first vaccination scheme and population?
+#' @param BaseLineModelsNames A character string of baseline model names.
+#' @param FileName A character string to use as the file name for the results.
+#' Defaults to ModelResults.
+#' @param SubFiles Numeric, defaults to 4. The number of subfiles to split the results into. Results are
+#' saved as \code{ModelResults1} etc. This must be smaller or equal to the number of models to be fitted.
+#' @param Folder Character string to use as the folder name for the results. 
+#' Defaults to "results"
+#' @param Path Character string specifying the path to the results folder. 
+#' Defaults to ".".
+#' @param ReRunAnalysis Logical, default to \code{FALSE}. Should the analysis be rerun.
+#' @param verbose Logical defaults to \code{FALSE}. Should progress messages be printed.
+#' @param Cores Numeric, defaults to the number of cores available minus 1.
+#'  How many cores to use for the analysis.
+#' @param Iterations Numeric, the number of iterations to use for the MCMC sampler. Defaults to 
+#' 10000.
+#' @param BurnIn Numeric, the number of iterations to use for the burn in. Defaults to 5000.
+#' @param Alpha Numeric, set the adaptive alpha value for the MCMC. Defaults to 0.99
+#' @param Treedepth Numeric, set the maximum tree depth. Defaults to 10.
+#' @param UpdateAlpha Numeric, set an updated alpha for models that have failed to fit.
+#' Defaults to 0.99999.
+#' @param HandleDLLs Logical, defaults to \code{FALSE}. Enables tracking and unloading of Stan DLL 
+#' objects. If not reachng R's DLL cap then this is not required.
+#' @param UpdateTreedepth Numeric, set the updated maximum tree depth for models that have failed to fit.
+#' @seealso clear_user_DLL Base_Model summarise_model_fit
+#' @return A dataframe of summarised results for the models specified to be fitted in the input dataframe.
+#' @export
+#' @importFrom parallel detectCores
+#' @importFrom dplyr select mutate filter do inner_join ungroup rename bind_rows bind_cols rowwise
[…]

### 2019-03-04 R/format_est_sd.R

@@ -1,0 +1,26 @@
+#' Format estimate and standard deviation
+#' 
+#' @param Estimate Numeric, the effect estimate. 
+#' @param SE Numeric, the standard error
+#' @param Digits Numeric, the number of digits to round to.
+#' @param Scientific Logical, defaults to \code{FALSE}. Should the results be
+#' displayed in scientific format.
+#'
+#' @return A format effect estimate with standard errors rounded to the specified digits.
+#' @export
+#' @examples
+#' 
+format_est_sd = function(Estimate, SE, Digits, Scientific = FALSE) {
+  
+  SE <-  SE %>%
+    signif(digits = Digits) %>% 
+    format(scientific = Scientific, digits = Digits)
+  
+  Estimate <- Estimate %>%
+    signif(digits = Digits) %>% 
+    format(scientific = Scientific, digits = Digits)
+  
+  EstSd <- paste0(Estimate, ' (', SE, ')') 
+  
+  return(EstSd)
+}

### 2019-03-04 R/format_model_summaries.R

@@ -1,0 +1,41 @@
+#' Summarise models
+#'
+#' @description Summarise fitted models and bind to descriptive information.
+#' @param Models A dataframe of fitted models with summary information
+#' @param ModelSpec A dataframe containing descriptive information about each model
+#' @param sep A character vector indicating the seperator used between the upper and lower credible intervals.
+#'  The default is ', '.
+#' @return A dataframe of formatted model summaries, including updated descriptive information.
+#' @export
+#'
+#' @importFrom dplyr mutate ungroup rowwise inner_join select
+#' @importFrom prettypublisher pretty_ci
+#' @examples
+#' 
+format_model_summaries = function(Models, ModelSpec, sep = ", ") {
+  
+  ## Assess Rhat convergance
+  Models <- Models %>% 
+    rowwise %>% 
+    mutate(Convergance = ifelse(length(Warnings) == 0, 'Yes', 'No')) %>% 
+    ungroup
+
+  ## Update model descriptions if updated since last fit
+  Models <- Models %>% 
+    inner_join(ModelSpec, by = c("Model"))
+  
+  ## Clean up effect sizes
+  Models <- Models %>%
+    rowwise %>% 
+    mutate(Effect = Effect %>% 
+             strsplit(split = '/') %>% 
+             unlist %>% 
+             pretty_ci(string = TRUE, sep = sep)) %>% 
+    ungroup
+  
+  
+  ## Remove spurious effects for models that do not include the change in policy
+  Models <- Models %>% 
+    mutate(Effect = Effect %>% replace(!grepl('PolicyChange', paste0(Predictors)), pretty_ci(0, 0, 0, sep = sep)))
+}
+

### 2019-03-04 R/justify_char_numeric.R

@@ -1,0 +1,23 @@
+#' Function to build justification framework for pander
+#'
+#' @param ColNum Numeric, the total number of columns
+#' @param Character Numeric vector, the numeric position of character columns.
+#' @param Numeric Numeric vector, the numeric position of numeric columns.
+#'
+#' @return A character string specifying the justification for a pander table.
+#' @export
+#'
+#' @examples
+#' 
+justify_char_numeric = function(ColNum = NULL, Character = NULL, Numeric = NULL) {
+  
+  JustPos <- seq(1, ColNum)
+  
+  JustPos <- JustPos %>% 
+    replace(JustPos %in% Character, 'l') %>%
+    replace(JustPos %in% Numeric, 'r') %>% 
+    replace(. %in% seq(1, ColNum), 'c') %>% 
+    paste0(collapse  = '')
+  
+  return(JustPos)
+}

### 2019-03-04 R/model_comparision.R

@@ -1,0 +1,48 @@
+
+#' Make Model Comparision, Filters by Scheme, Country of Birth, and Convergance
+#'
+#' @param DF A dataframe of summarised model results.
+#' @param Scheme A character string specifying the scheme to select models for.
+#' @param Pop A character string speccifying the population to select models for.
+#' @param Digits Numeric, the number of digits to round to.
+#'
+#' @seealso format_est_sd
+#' @return A summarised table comparing model results
+#' @export
+#' @importFrom dplyr filter mutate select rename arrange
+#' @importFrom prettypublisher pretty_round
+#' @importFrom purrr map_lgl
+#' @examples
+#' 
+model_comparison = function(DF, Scheme, Pop, Digits = 3) {
+ 
+  ##Sumarise model components
+  ComTable <- DF %>% 
+    filter(VacScheme %in% Scheme) %>% 
+    filter(CoB %in% Pop) %>%
+    filter(Convergance %in% 'Yes') %>%
+    mutate(`Policy Change` = ifelse(grepl('PolicyChange', Predictors), 'Yes', 'No')) %>% 
+    mutate(`Age` = ifelse(grepl('YrsFollowUp', Predictors), 'Yes', 'No')) %>%
+    mutate(`Year of study entry` = ifelse(grepl('YearEligib', Predictors), 'Yes', 'No')) %>%
+    mutate(`UK born rates` = ifelse((grepl('UKRate', Predictors, fixed = TRUE) &  !grepl('NonUKRate', Predictors, fixed = TRUE)) |
+                                      map_lgl(Predictors, ~sum(grepl('UKRate', ., fixed = TRUE)) == 2) |
+                                    (CoB %in% 'UK born' & grepl('TotalRate', Predictors)), 'Yes', 'No')) %>% 
+    mutate(`Non-UK born rates` = ifelse(grepl('NonUKRate', Predictors) | (CoB %in% 'Non-UK born' & grepl('TotalRate', Predictors)), 'Yes', 'No'))
+  
+  ## Arrange and order by fit stats 
+  ComTable <-  ComTable %>% 
[…]

### 2019-03-04 R/pretty_round_inf.R

@@ -1,0 +1,19 @@
+#' Pretty format after checking a result is smaller than a limit
+#'
+#' @param x Numeric or character effect size estimate
+#' @param digits Numeric defaults to 2. The number of digits to round to.
+#' @param ... Additional arguements to pass to \code{OR_inf_rep}.
+#'
+#' @return A prettty effect estimate bounded by some maximum size.
+#' @export
+#' @importFrom prettypublisher pretty_round
+#'
+#' @examples
+#' 
+pretty_round_inf = function(x, digits = 2, ...) {
+  x <- OR_inf_rep(x,  ...)
+  x <- pretty_round(x, digits = digits)
+  return(x)
+}
+
+

### 2019-03-04 R/sim_cohort_vac_cam.R

@@ -1,0 +1,39 @@
+
+#' Identify those that are eligible for a vaccination program and track age specific incidence rates.
+#'
+#' @description This function identifies those that are eligible for vaccination programs and tracks their age
+#' specific incidence over years since vaccination. It does this by,
+#' - Produces a data set with year of eligibility as a variable
+#' - Does this for both the universal and targeted schemes
+#' - Adds a variables for years since entry to scheme - both universal and targetted
+#' - Adds a flag for before or after policy change (i.e 2005 switch)
+#'
+#' @param df A dataframe
+#' @param eligib_vector Numeric year identifying eligibility.
+#' @param campaign A character string identifying the name of the campaign.
+#' @param policy_change_yr Numeric year identifying the year of the policy change.
+#'
+#' @return A simulate cohort for a given vaccination program correctly formatted as a
+#' retrospective cohort.
+#' @export
+#' @importFrom dplyr mutate filter
+#' @examples
+#'
+sim_cohort_vac_cam = function(df, eligib_vector, campaign, policy_change_yr = 2005) {
+
+ tmp <-  df %>%
+    filter(!(Age %in% c('All cases (crude)', 'All cases (adj)')), !(Year %in% '2016')) %>%
+    mutate(YrsFollowUp = Age %>%
+             as.character %>%
+             replace(Age %in% '90+', '90') %>%
+             as.numeric) %>%
+    mutate(YrsFollowUp = YrsFollowUp - eligib_vector) %>%
+    mutate(YrsFollowUp = ifelse(YrsFollowUp < 0, NA, YrsFollowUp)) %>%
+    mutate(YearEligib = as.numeric(Year) - YrsFollowUp) %>%
+    mutate(YearEligib = ifelse(YearEligib < min(as.numeric(Year)), NA, YearEligib)) %>%
+    mutate(PolicyChange = ifelse(YearEligib < policy_change_yr, 'Pre change', 'Post change') %>% as.factor) %>%
+    mutate(VacScheme = campaign) %>%
+    filter(!is.na(YrsFollowUp), !(is.na(YearEligib)))
+
+  return(tmp)
+}

### 2019-03-04 R/store_warnings.R

@@ -1,0 +1,16 @@
+
+#' Detect Warnings and Store as a List
+#' @param ListFits List of fitted model objects
+#'
+#' @return A list of model warnings
+#' @export
+#' @importFrom dplyr nth
+#' @importFrom purrr map transpose quietly
+#' @examples
+#' 
+store_warnings = function(ListFits) {
+  ListFits %>% 
+    map(quietly(function(x) summary(x))) %>%
+    transpose %>% 
+    nth(n = 3) 
+}

### 2019-03-04 R/summarise_model_fit.R

@@ -1,0 +1,65 @@
+#' Summarise model fit results  
+#'
+#' @param ModelsWithFits Dataframe containg fitted model objects (stored as a Fit variable).
+#' @param cores Numeric, the number of cores to use for parralisation (used by loo). Default
+#' is 1.
+#' @return A datafrmae containing summarised model fit statistics.
+#' @export
+#' @importFrom dplyr bind_cols do rowwise mutate_each select mutate_all slice first nth ungroup mutate
+#' @importFrom purrr map
+#' @importFrom broom tidy
+#' @importFrom brms LOO WAIC
+#' @examples
+#' 
+summarise_model_fit = function(ModelsWithFits, Cores = 1) {
+  
+  ## Summarise model results
+  Models <- ModelsWithFits %>% 
+    bind_cols(ModelsWithFits  %>% 
+                rowwise %>% 
+                do(
+                   TidySum = .$Fit %>% 
+                     tidy(conf.int = TRUE, conf.level = 0.95) %>% 
+                     mutate_at(.vars = vars(estimate, lower, upper), .funs = funs(exp(.))), 
+                   Effect = .$Fit %>% 
+                     tidy(conf.int = TRUE, conf.level = 0.95) %>% 
+                     select(estimate, lower, upper) %>%
+                     mutate_all(funs(exp(.))) %>% 
+                     slice(2) %>% 
+                     unlist(., use.names = FALSE) %>%
+                     paste(collapse = '/'),
+                   WAIC_ob = .$Fit %>% WAIC,
+                   LOOIC_ob = .$Fit %>% LOO(reloo = FALSE, cores = Cores))
+    ) %>% 
+    mutate(LOOIC = LOOIC_ob %>% map(~.$estimates[3, 1]) %>% unlist,
+           LOOICSE = LOOIC_ob %>% map(~.$estimates[3, 2]) %>% unlist,
+           WAIC = WAIC_ob %>% map(~.$estimates[3, 1]) %>% unlist,
+           WAICSE = WAIC_ob %>% map(~.$estimates[3, 2]) %>% unlist
+           
+    )
+  
+  
+  ## Add computed log pointwise predictive density
+  ## Also add estimated effective number of parameters
+  Models <- Models %>% 
+    mutate(elpd = map_dbl(WAIC_ob, ~.$estimates[1, 1]),
[…]

### 2019-03-04 data-raw/create_package.R

@@ -1,0 +1,28 @@
+
+#install.packages("devtools")
+library(devtools)
+
+# init stuffs
+use_data_raw()
+use_cran_badge()
+use_mit_license()
+use_readme_rmd()
+use_testthat()
+use_roxygen_md()
+
+
+#Suggest
+use_package("pkgdown", type = "Suggests")
+use_package("devtools", type = "Suggests")
+
+#Import
+
+
+#Vignettes
+use_vignette("paper")
+
+##Build site, and make pkgdown
+devtools::document()
+devtools::build_vignettes()
+devtools::build()
+pkgdown::build_site()

### 2019-03-04 data-raw/extract_not_sum_data.R

@@ -1,0 +1,165 @@
+## All cleaning is done via the tbinenglanddataclean package see help files for details
+## See the package website: http://www.samabbott.co.uk/tbinenglanddataclean for details of data sources
+## Raw data needs to be stored in the data-raw folder
+library(tidyverse)
+library(tbinenglanddataclean)
+library(prettypublisher)
+library(scales)
+library(h2o)
+
+safe_ets_clean <- purrr::safely(clean_munge_ets_2016)
+
+ets <- safe_ets_clean(data_path = "tb_data/ETS/ETS_2016_extract/SAbbott_BCG dataset_version2_final140217.dta",
+                     return = TRUE,
+                     save = TRUE,
+                     save_name = "clean_ets_2016",
+                     save_path = "tb_data/tbinenglanddataclean",
+                     save_format = c("rds", "csv"),
+                     verbose = FALSE)$result
+
+if (!is.null(ets)) {
+  ## Evaluate missing data
+  ets %>% 
+    filter(year %in% 2000:2015) %>% 
+    select(ukborn, year, age) %>% 
+    map_dbl(~sum(is.na(.))/length(.))
+  
+  ## Missing data is only an obvious issue for UK birth status with around 7% missing.
+  ## Does the missing data change over time
+  
+  ## Look at missing data by year
+  ets %>% 
+    count(ukborn, year) %>% 
+    add_count(year, wt = n) %>% 
+    mutate(prop = n / nn) %>% 
+    ggplot(aes(x = year, y = prop, col = ukborn)) +
+    scale_y_continuous(labels = percent, limits = c(0,NA), breaks = seq(0, 1, 0.1)) +
+    geom_point(size = 1.2) +
+    geom_line(alpha = 0.8, size = 1.1) +
+    theme_minimal() +
+    theme(legend.position = "top") +
+    facet_wrap(~ukborn, nrow =1, scales = "free_y")
+  
+  ## Yes missing data goes down over time, with increasing cases
+  ## in both the non-UK born and UK born over this period.
+  ## Therefore we need to impute - do this using a nonlinear algorithm including variables reasonably
+  ## related to missingness
+  
+  ## Split data into complete UKborn status and not complete groups
+  complete_ets <- ets %>% 
+    drop_na(ukborn)
+  
+  incomplete_ets <- ets %>% 
+    filter(is.na(ukborn))
+  
+  ## Launch h2o
+  h2o.init(min_mem_size = "14G",
[…]

### 2019-03-04 data-raw/get_incidence_rates.R

@@ -1,0 +1,105 @@
+## All cleaning is done via the tbinenglanddataclean package see help files for details
+## See the package website: http://www.samabbott.co.uk/tbinenglanddataclean for details of data sources
+## Raw data needs to be stored in the data-raw folder
+library(tidyverse)
+library(tbinenglanddataclean)
+
+if (dir.exists("tb_data")) {
+  clean_munge_ets_2016(data_path = "tb_data/ETS/ETS_2016_extract/SAbbott_BCG dataset_version2_final140217.dta",
+                       return = FALSE,
+                       save = TRUE,
+                       save_name = "clean_ets_2016",
+                       save_path = "tb_data/tbinenglanddataclean",
+                       save_format = c("rds", "csv"),
+                       verbose = FALSE)
+  
+  
+  ## Set up demographic data if required (or run tbinengladdataclean vignette on your local computer)
+  ## ONS
+  clean_demographics_uk(data_path = "tb_data/UK_demographics",
+                        demo_2000 = "UK_2000_age.csv",
+                        demo_2001_2015 = "UK_2001_2015_age.csv",
+                        countries = c("E"),
+                        return = FALSE,
+                        save = TRUE,
+                        save_name = "E_demo_2000_2015",
+                        save_path = "tb_data/tbinenglanddataclean",
+                        save_format = c("rds", "csv"),
+                        verbose = FALSE)
+  
+  
+  
+  ## LFS
+  clean_labour_force_survey(data_path = "tb_data/LFS",
+                            return = FALSE,
+                            save = TRUE,
+                            save_name = "formatted_LFS_2000_2016",
+                            save_path = "tb_data/tbinenglanddataclean",
+                            save_format = c("rds", "csv"),
+                            verbose = FALSE)
+  
+  
+  
+  ## Combined ONS, LFS data
+  combine_ons_with_lfs(data_path = "tb_data/tbinenglanddataclean",
+                       ons_name = "E_demo_2000_2015.rds",
+                       lfs_name = "formatted_LFS_2000_2016.rds",
+                       countries = "England",
+                       return = FALSE,
+                       save = TRUE,
+                       save_name = "E_ons_lfs_2000_2016",
+                       save_path = "tb_data/tbinenglanddataclean",
+                       save_format = c("rds", "csv"),
+                       verbose = FALSE) 
+  
+  
+  
+  ## Calculate incidence and store in the repository folder - note only checks for incidence data
[…]

### 2019-03-04 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(DIrectEffBCGPolicyChange)
+
+test_check("DIrectEffBCGPolicyChange")

### 2019-03-04 vignettes/scripts/descriptive_data_sum.R

@@ -1,0 +1,165 @@
+
+TotalStudyCases <- StudyCasesSum$n %>% sum
+
+## Percentage of cases with complete UK birth status
+CompleteBirthStatus <- StudyCasesSum %>% 
+  filter(!is.na(ukborn)) %>%
+  select(n) %>%
+  sum
+CompleteBirthStatusPer <- CompleteBirthStatus %>% 
+  pretty_percentage(TotalStudyCases, digits = 0)
+
+##Cases that were UK born
+CasesUKborn <- StudyCasesSum %>% 
+  filter(ukborn %in% 'UK Born') %>%
+  select(n) %>% 
+  sum
+CasesUKbornPer <- CasesUKborn %>% 
+  pretty_percentage(CompleteBirthStatus, digits = 0)
+
+##Eligible For schemes
+eligible_filter <-  function(df, cob_filter = NULL) {
+  df %>% 
+    filter(CoB %in% cob_filter) %>%
+      group_by(VacScheme, PolicyChange) %>% 
+      summarise(Cases = sum(Cases), 
+                Population = sum(Population), 
+                Incidence = epi_conf(Cases, Population)$est * RateScale,
+                Inc_LCI = epi_conf(Cases, Population)$lower * RateScale,
+                Inc_UCI = epi_conf(Cases, Population)$upper * RateScale,
+                PrettyInc = pretty_ci(Incidence, 
+                                      lci = Inc_LCI, 
+                                      uci = Inc_UCI, 
+                                      inline = TRUE,
+                                      note = "per 100,000 (95%CI ",
+                                      replace_bracket = FALSE)) %>% 
+      ungroup
+    
+}
+
+EligibForSchemes <- RestCohortVacInc %>%
+  eligible_filter("UK born")
+
+CasesUniPrePost <- EligibForSchemes %>% 
+  filter(VacScheme %in% 'Universal school-age (14)') %>%
+  pull(Cases) %>%
+  purrr::set_names(c("Pre change", "Post change"))
+
+CasesUni <- CasesUniPrePost %>% 
+  sum
+
+CasesUniPer <- CasesUni %>% 
+  pretty_percentage(CasesUKborn, digits = 0)
+
+IncUni <- EligibForSchemes %>% 
+            filter(VacScheme %in% 'Universal school-age (14)') %>% 
[…]

### 2019-03-04 vignettes/scripts/exploratory_plots.R

@@ -1,0 +1,142 @@
+## Look at incidence rates over time in the general population, and those effected directly by the policy change
+GraphStudySpecIncRates <- StudySpecIncRates %>%
+  mutate(`Policy change (2005)` = ifelse(Year < 2005, 'Pre change', 'Post change') %>% 
+           factor(levels = c('Pre change', 'Post change'))) %>% 
+  mutate(CoB = CoB %>% relevel("UK born")) %>% 
+  ggplot(aes(
+    x = Year,
+    y = Incidence,
+    colour = `Policy change (2005)`,
+    fill = `Policy change (2005)`
+  )) +
+  geom_pointrange(aes(ymin = Inc_LCI, ymax = Inc_UCI), fatten = 1.1, size = 1) +
+  geom_line(aes(group = `Policy change (2005)`), size = 0.8, alpha = 0.6) +
+  geom_vline(xintercept = 2004) +
+  scale_color_brewer(palette = "Dark2") +
+  labs(y = "Incidence rate (per 100,000)") +
+  PaperTheme() +
+  
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  facet_grid(CoB ~ `Age group`, scale = 'free_y') +
+  theme(legend.position = "top")
+
+##  Look at average in cohorts between those Pre and post policy change
+##  Look at average in cohorts between those Pre and post policy change
+## define the scale of incidence rates
+AvgRestCohortVacInc <- RestCohortVacInc %>%
+  group_by(VacScheme, PolicyChange, Age, CoB) %>%                                                       
+  summarise(
+    Incidence = epi_conf(Cases, Population)$est * RateScale,
+    LCI = epi_conf(Cases, Population)$lower * RateScale,
+    UCI = epi_conf(Cases, Population)$upper * RateScale
+  )  %>%
+  ungroup %>% 
+  mutate(Cohort = case_when(VacScheme %in% "Universal school-age (14)" 
+                            & CoB %in% "UK born" &
+                              PolicyChange %in% "Pre change" ~ "UNCUK14",
+                            VacScheme %in% "Universal school-age (14)" 
+                            & CoB %in% "UK born" &
+                              PolicyChange %in% "Post change" ~ "UCUK14",
+                            VacScheme %in% "Targeted high-risk neonates (0)" 
+                            & CoB %in% "UK born" &
+                              PolicyChange %in% "Pre change" ~ "TNCUKBirth",
[…]

### 2019-03-04 vignettes/scripts/packages.R

@@ -1,0 +1,44 @@
+
+# packages used in this module --------------------------------------------
+package_names <- c('tidyverse', 'knitr', 'pander', 'haven', 'epiR', 
+                   'scales', 'broom', 'shinystan', 'brms', 'pomp', 'stringr', 
+                   'prettypublisher', 'tbinenglanddataclean')
+
+dev_tools <- rep(NA, length(package_names))
+names(dev_tools) <- package_names
+
+dev_tools["prettypublisher"] <- "seabbs/prettypublisher"
+dev_tools["tbinenglanddataclean"] <- "seabbs/tbinenglanddataclean"
+
+#change to repo url if dev tools are required for a package
+
+##Load a single package
+pkgLoad_single <- function(x, y)
+{
+  if (!require(x,character.only = TRUE))
+  {
+    if (is.na(y))
+    {
+      install.packages(x,dep = TRUE, repos = 'http://www.stats.bris.ac.uk/R/')
+    }else{
+      install_github(y)
+    }
+    if (!require(x,character.only = TRUE)) stop("Package not found")
+  }
+  return(paste0('Loaded ', x, ' successfully'))
+}
+
+## Append devtools and load all packages
+pkgLoad <- function(pck_names, pck_dev)
+{
+  pck_names <- c('devtools', pck_names)
+  pck_dev <- c(NA, pck_dev)
+  pck <- mapply(pkgLoad_single, pck_names, pck_dev)
+  return(pck)
+}
+
+## Execute and load all packages
+pck <- pkgLoad(package_names, dev_tools)
+print(pck)
+
+

### 2019-03-04 vignettes/scripts/restrict_cohorts_study_crit.R

@@ -1,0 +1,93 @@
+## Restrict data to those years with 5 complete years of follow up (i.e post 2010) and standardised results for 5 years of follow up
+RestCohortVacInc <- CohortVacInc %>% 
+  filter(YearEligib <= 2010) %>%
+  filter(YrsFollowUp <= 5)
+
+## Estimate total population values
+cases_demo_incidence_total <- cases_demo_incidence %>% 
+  group_by(Year, CoB) %>% 
+  summarise(TotalCases = sum(Cases, na.rm = TRUE), 
+            TotalPopulation = sum(Population, na.rm = TRUE)) 
+
+
+## Add total cases and total population for both Non-UK born and UK-born
+cases_demo_incidence_uk_strat <- cases_demo_incidence_total %>%
+  spread(CoB, TotalCases) %>% 
+  select(Year, `Non-UK born`, `UK born`) %>%
+  rename(NonUKCases = `Non-UK born`, UKCases = `UK born`) %>% 
+  group_by(Year) %>% 
+  summarise(NonUKCases = sum(NonUKCases, na.rm = TRUE), 
+            UKCases = sum(UKCases, na.rm = TRUE)) %>% 
+  full_join(cases_demo_incidence_total %>%
+              spread(CoB, TotalPopulation) %>% 
+              select(Year, `Non-UK born`, `UK born`) %>% 
+              rename(NonUKPop = `Non-UK born`, UKPop = `UK born`) %>% 
+              group_by(Year) %>% 
+              summarise(NonUKPop = sum(NonUKPop, na.rm = TRUE), 
+                        UKPop = sum(UKPop, na.rm = TRUE)), 
+            key = Year
+  )
+
+## Bind UK born and non UK born rates back into study populations
+## Exclude cases and population in the study population of interest
+RestCohortVacInc <- RestCohortVacInc %>% 
+  left_join(cases_demo_incidence_uk_strat, key = c(Year)) %>% 
+  filter(!is.na(VacScheme)) %>% 
+  mutate(UKCases = case_when(CoB %in% "UK born" ~ UKCases - Cases,
+                             TRUE ~ UKCases),
+         UKPop = case_when(CoB %in% "UK born" ~ UKPop - Population,
+                           TRUE ~ UKPop),
+         NonUKCases = case_when(CoB %in% "Non-UK born" ~ NonUKCases - Cases,
+                                TRUE ~ NonUKCases),
[…]

### 2019-03-04 README.Rmd
added badges and bumped R version

@@
 
-**Background**
+[![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/DirectEffBCGPolicyChange)
+[![badge](https://img.shields.io/badge/Launch-Analysis-lightblue.svg)](https://mybinder.org/v2/gh/seabbs/DirectEffBCGPolicyChange/master?urlpath=rstudio)
+[![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)
+
+## Abstract
+
+### Background
 
@@
 
-**Methods**
+### Methods
 
@@
 
-**Results:**
+### Results
 
@@
 
-**Conclusions**
+### Conclusions
 
@@
 
-# Docker
+## Reproducibility
+
+## Docker
 

### 2019-03-04 README.Rmd
reproducibility sweep and DOI

@@
 
-## Docker
+### Repository structure
+
+The repository is structured as an R package. It has the following structure:
+
+- **data-raw:** Raw data processing.
+- **data:** Processed data.
+- **R:** Supporting R functions
+- **vignettes:** Analysis paper and results.
+
+### Manual install
+
+- Install R (analysis run with `3.5.2`) and Rstudio.
+
+- Download the analysis folder from [https://github.com/seabbs/DirectEffBCGPolicyChange/archive/master.zip](https://github.com/seabbs/DirectEffBCGPolicyChange/archive/master.zip) or use `git clone`, as follows, in the command line (not the R terminal).
+
+```{bash clone-package, eval = FALSE}
+git clone https://github.com/seabbs/DirectEffBCGPolicyChange.git
+```
+
+- Once this has been downloaded click on the project file (`DirectEffBCGPolicyChange.Rproj`).
+
+- Install the analysis dependencies and build the package using the following.
+To enable more robust reproducibility consider using the [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/index.html) package versioned locked to R `3.5.2`.
+
+```{r get-packages, eval = FALSE}
+
+#install.packages("devtools")
+
+# To build locally
+devtools::install_dev_deps(dependencies = TRUE)
+devtools::install()
+
+# Alternatively to remote install
+devtools::install_github("seabbs/DirectEffBCGPolicyChange", dependencies = TRUE)
+```
+
+- Load the analysis results by running `vignettes/paper.Rmd`.
+The complete analysis is not reproducible as the raw data cannot be published, however all interim results are stored in the repository (`data`) and these can be explored.
+
+- Alternatively the complete analysis (along with documentation) can be reconstructed using `make` in the project root directory.
+
+### Docker
 
@@
 ```
+
+Alternatively the analysis environment can be accessed via [binder](https://mybinder.org/v2/gh/seabbs/DirectEffBCGPolicyChange/master?urlpath=rstudio).

### 2019-03-04 README.Rmd
readme updates

@@
 
-### Methods
+#### Methods
 
@@
 
-The change in BCG policy was associated with reduced incidence of TB in the study population.
+Withdrawing universal vaccination at school-age and targeting BCG vaccination towards high-risk neonates was associated with reduced incidence of TB in England.
 This was largely driven by reductions in the non-UK born.
-There was a slight increase in those at school-age, who were UK born.
+There was a slight increase in UK born school-age cases.
 

### 2019-03-04 README.Rmd
submission hurdle

@@
 [![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)
+
+[Sam Abbott](https://www.samabbott.co.uk), Hannah Christensen, Nicky Welton, Ellen Brooks-Pollock
 
@@
 - **vignettes:** Analysis paper and results.
+- **peer-review** Documentation required for peer review.
 
@@
 
-- Install R (analysis run with `3.5.2`) and Rstudio.
+- Install R (analysis run with `3.5.2`) and Rstudio (altnernatively use Docker as outlined below).
 

### 2019-03-04 peer-review/cover_letter.Rmd
submission hurdle

---
title: ""
output:
   html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editor,

Please find attached our manuscript, entitled "Estimating the effect of the 2005 change in BCG policy in England: A retrospective cohort study".

We feel our manuscript will be of interest to you and the journal's readership because Bacillus Calmette-Guérin (BCG) is one of the most widely-used vaccines worldwide and our study explores the impact of changing BCG vaccination policy. We found evidence to suggest that withdrawing universal vaccination at school-age and targeting BCG vaccination towards high-risk neonates was associated with reduced incidence of Tuberculosis (TB) in England. These findings provide support for current World Health Organization recommendations. We further found that the reduced incidence in TB was largely driven by reductions in the non-UK born, with a slight increase in UK born school-age cases. Our study benefits from the use of data from the Enhanced Tuberculosis surveillance system, which contains demographic data on active TB cases in England. We made use of Poisson and Negative Binomial regression models, with imputed UK birth status, in order to estimate the impact of the change in policy on TB.

We do not have any issues relating to the journal policies. All authors have approved this version of the manuscript for submission and the content of the manuscript has not been published or submitted elsewhere.

With kind regards, on behalf of all the co-authors,

Sam Abbott

### 2019-03-04 README.Rmd
readme improvements

--- 2019-03-04 README.Rmd
@@
 
-- **data-raw:** Raw data processing.
-- **data:** Processed data.
-- **R:** Supporting R functions
-- **vignettes:** Analysis paper and results.
+- `data-raw`: Raw data processing.
+- `data`: Processed data.
+- `R`: Supporting R functions.
+- `vignettes`: Analysis paper and results.
+- `peer-review`: Documentation required for peer review.
 
@@
 
-- Install R (analysis run with `3.5.2`) and Rstudio.
+- Install R (analysis run with `3.5.2`) and Rstudio (alternatively use Docker as outlined below).
 
@@
 The complete analysis is not reproducible as the raw data cannot be published, however all interim results are stored in the repository (`data`) and these can be explored.
-
-- Alternatively the complete analysis (along with documentation) can be reconstructed using `make` in the project root directory.
+Alternatively the complete analysis (along with documentation) can be reconstructed using `make` in the project root directory.
 

### 2019-03-05 README.Rmd
roll back to R 3.5.0

@@
 
-- Install R (analysis run with `3.5.2`) and Rstudio (alternatively use Docker as outlined below).
+- Install R (analysis run with `3.5.0`) and Rstudio (alternatively use Docker as outlined below).
 
@@
 - Install the analysis dependencies and build the package using the following.
-To enable more robust reproducibility consider using the [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/index.html) package versioned locked to R `3.5.2`.
+To enable more robust reproducibility consider using the [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/index.html) package versioned locked to R `3.5.0`.
 

### 2019-03-05 README.Rmd
doc refresh

@@
 
+[![badge](https://img.shields.io/badge/Launch-Analysis-lightblue.svg)](https://mybinder.org/v2/gh/seabbs/DirectEffBCGPolicyChange/master?urlpath=rstudio)
 [![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/DirectEffBCGPolicyChange)
-[![badge](https://img.shields.io/badge/Launch-Analysis-lightblue.svg)](https://mybinder.org/v2/gh/seabbs/DirectEffBCGPolicyChange/master?urlpath=rstudio)
 [![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)

### 2019-04-01 R/model_comparison.R
prep for eurosurveillance submission

@@ -1,5 +1,5 @@
 
-#' Make Model Comparision, Filters by Scheme, Country of Birth, and Convergance
+#' Make Model Comparison, Filters by Scheme, Country of Birth, and Convergance
 #'
 #' @param DF A dataframe of summarised model results.
 #' @param Scheme A character string specifying the scheme to select models for.

### 2019-04-01 README.Rmd

@@
 [![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/DirectEffBCGPolicyChange)
+[![Paper](https://img.shields.io/badge/Paper-10.1101/567511-lightgreen.svg)](https://doi.org/10.1101/567511)
 [![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)

### 2019-04-01 R/utils-pipe.R
added pass td pass through saving for formated results -> thesis

@@ -9,3 +9,14 @@
 #' @importFrom magrittr %>%
 #' @usage lhs \%>\% rhs
 NULL
+#' Pass through pipe operator
+#'
+#' See \code{magrittr::\link[magrittr]{\%T>\%}} for details.
+#'
+#' @name %T>%
+#' @rdname pass-pipe
+#' @keywords external
+#' @export
+#' @importFrom magrittr %T>%
+#' @usage lhs \%T>\% print
+NULL

### 2019-10-07 peer-review/review-response.Rmd
added reviewer comments

---
title: "Reviewer Response"
output:
  word_document: default
  html_document:
    df_print: paged
---

Dear Editor,

Thank you for sending review reports for our manuscript. The reviewer’s comments were useful and have greatly improved the paper.

We enclose a revised version of the manuscript (clean and with tracked changes), which addresses the reviewer’s comments and suggestions. The reviewer’s comments are shown below with our response to each comment in italics.

Yours sincerely,

Sam Abbott

## Authors' response to Editorial Requests

GENERAL

- In both the abstract and the main text please define all abbreviations at first use (e.g. UK; gradient boosting method – is this GBM?; 95%CI – is CI confidence interval? Credible interval?; MNAR in the discussion – is this missing not at random mechanism?). For Tables and Figures, please define all abbreviations used in a footnote below the Figure or Table.

- If possible, please use square brackets for the references with the punctuation following the reference.

- Tables should not have any empty cells as design element or because information is not available ('NA' (subsequently defined as not available or not applicable) can be used for example).

- For all Figure and Table titles, when possible, please indicate at the end of the title, the geographical area, period and total number of items covered by the illustration. E.g. Table 1: Summary of relevance and eligibility criteria for each cohort, United Kingdom, 2000-2010 (n = xxx individuals)

- Please check the overall text versus the tables for the name of the intervals (e.g. 95% confidence intervals vs 95% credible intervals). To give an example: In the abstract and in the main text subsection called ‘Adjusted estimates of the effects of the change in policy on school-age children’ there is ‘IRR: 0.74 (95%CI 0.61, 0.88)’ but in the table 2 there is ‘0.74 (0.61, 0.88)’ defined as IRR (95% CrI) CrI: Credible interval.

ABSTRACT

- While keeping the word count below 250 words, please add an ‘aim’ subsection, describing briefly the aim of the paper, and please specify the study type (e.g. observational study?) also in the abstract (either in the aim or the methods).

METHODS

Data sources 
- Does the Labour Force Survey record only the number of adults of working age (employed or unemployed) in the country, or does it record more people than these (e.g. their family members such as children)? Please describe the population and demographics recorded in the Labour Force Survey in more detail.

Statistical methods

- While it is mentioned in the methods that 95% confidence intervals were calculated around the IR values, it is not described what the intervals are around the IRRs. Please check the earlier editorial comment about the intervals that are presented in text and tables around the IRRs.

Ethical statement

- An ethical statement should be included (https://www.eurosurveillance.org/for-authors#ethical%20statement); if ethical approval was not needed, this should be stated in the manuscript. In this case, if possible, it should be please explained why.

RESULTS

= ‘Of notifications with a known birth status 27% (29096/106765) were UK born, in comparison to 33% (2634/8055) in cases with an imputed birth status.’ Just for editor’s understanding: does this mean that ‘Among notifications with a known birth status, 27% had the status ‘born in the UK’, while among notifications with an imputed birth status, 33% had the status ‘born in the UK’?

## Authors' response to Reviewer #1:

This is an interesting paper in which the authors address a key TB control issue which is measuring the impact of BCG policy change in the incidence of tuberculosis in children. Whether and when targeted BCG vaccination should replace universal vaccination depends on several factors and may be a matter of discussion within national TB control programs, although in Europe most countries have a targeted BCG policy. The question is worthwhile investigating and discussing. I suggest publication in Eurosurveillance provided that the minor changes suggested are taken into account. HOWEVER, given the importance of the statistical methods in this analysis, which I do not feel confident to analyse, I recommend that a tough review of this part, and the results, be made by a statistician, which opinion is crucial to decide whether the paper is able to be published or not.

Introduction

1.	Paragraph 2. In France, universal vaccination was implemented in all children before entering "community care" and primary school at the latest, and not in neonates. 
Methods

1.	Paragraph 1. The geographical scale of TB local incidence is not specified and is an important piece of information that would be worth mentioning at this stage.

2.	Paragraph 2. Not clear whether the LFS which reports "employment circumstances" of the UK population reports also data in children, who are not employed. We assume the answer is yes which should be probably clarified.

3.	Definition of cohorts. Not totally clear to me. Example for comparison cohort 2: a child born in 2000 and who is diagnosed with TB at 2 years old (2002): the vaccination programme in 2002 is universal not targeted, as seen in Table 1. Maybe further clarify this crucial point.

Statistical Methods

This is a key point that should be reviewed by a statistician; I do not feel confident on giving my opinion on this.

Results

1.	No comments, except that I find some parts of the results difficult to understand for a "common" public health professional" with no deep knowledge on statistics and modelling methods.

Discussion

1.	BCG protects mainly against severe disease (meningitis). This point is not discussed at all. An additional analysis could have investigated the impact of policy change on the incidence of meningeal TB in children, which could have broaden the consequences of policy change in the epidemiology of TB in children. Maybe a word on this should be said in Introduction or Discussion.

2.	Other interventions may decrease TB incidence, such as a better screening, a change in screening policy in migrants, a better case management…This is not discussed. Could these potential interventions have also played a role in the results presented here?

3.	No figures of BCG vaccination coverage trends in targeted groups are given (except the 68% mentioned in the Discussion). This determines to a large extent the impact of vaccination. A brief description of this could be given here, or in Introduction.

4.	BCG has an impact mainly in severe forms of TB, and prevents about 50% of pulmonary TB in children. But children, to my knowledge, do not frequently transmit TB. So BCG has rather poor effect on TB transmission. The authors speak in their discussion on indirect effects and ongoing transmission. Is this really the case with BCG and TB?

## Authors' response to Reviewer #2:

Thank you for providing an excellent overview on the effects of the policy changes in BCG vaccination in England.

## Authors' response to Reviewer #3:

This study is of high relevance, not just for the England, but for the global tuberculosis community: The UK (England and Wales notably) had been one of the few countries with a universal vaccination of adolescents and a clinical trial demonstrating high efficacy for at least 20 years in the age vaccinated groups and beyond to include young adults during follow-up.

From the switch from universal adolescent vaccination to selective (largely the descendant newborn of foreign-born ancestors) one would thus hypothetically expect an increase in the incidence rate among the former and a decrease in the incidence rate among the latter (among the respective cohorts in both groups), if confounding resulting from secular trends could be properly controlled for. This study aimed to do precisely that.

The authors created 8 cohorts, each for a 5-year follow-up, using tuberculosis surveillance data for the numerator and Labor Force Survey data for the population denominator. These procedures are all clearly explained. As it gets substantially steeper and difficult to follow for the more general user's consumption is when the models are elaborated in the same paragraph. Perhaps it would help to split the elements of describing the basic epidemiology of incidence rates and ratios from the latter into two separate paragraphs, so that those unfamiliar with "negative binomial models", "penalised by model complexity", and the intricacies of Bayesian estimation with Markov Chain Monte Carlo, etc. This would clarify that here is a basic methodology, and here is a complex modeling description, the former for every epidemiologist, the latter only for the seasoned expert in statistical modeling using special techniques. I concur that it should and has to be there, but the more 
basic working techniques of calculation might be preferably "visually" separated from complex modeling techniques, notably when unexplained acronyms such as "GBM" are interspersed that do not help in easily following what the modeler actually did. Nor does it become that clear why exactly these models were chosen and not something else. The reader should probably have some inkling what the motivation for model selection is.

In any case, it becomes again clear that the procedures resulted in obtaining adjusted estimates of the effects of the policy change from universal adolescent to selective neonatal vaccination. The effect was measurable but seemingly modest. Apparently, the model required adjustments by including the change in policy, age, incidence rates in the non-UK born, and year of eligibility as a random effect. To the non-initiated, all these elements make intuitively sense and reflect the complexity of interplaying variables that all played a substantial role in what was finally observable. Such complexity may also contribute to variability and thus uncertainty about the unbiased estimation of the magnitude of the observable effect. However, the authors later also state that all models (does this refer only to the non-UK born?) that included at least age in the adjustment resulted in comparable effects. As parsimony is king in all modeling, perhaps it could be stated more 
explicitly that always the model with the minimum required variables was used and perhaps why recourse was taken to retaining more complex models if a more simple one gave actually comparable results. Perhaps I am missing here something, but then it would indeed be useful to review the sum of the modeling exercises to the point.

In table 3, among the other variables, is also the year of study eligibility (please correct spelling). The point estimate incidence rate ratio varies from 0.83 to 1.12 (UK born) with no discernible trend. Is this therefore an expression of the intrinsic uncertainty of the effect of the model? If so, then perhaps the authors would have to be more explicit that there remains substantial uncertainty in the modeled effect as small as it may be. Thus perhaps the variability in the estimated magnitude of effect for cases prevented in the following paragraph which is substantial is therefore not all that meaningful after all (CrIs not uncommonly overlapping zero).

Thus, in the end in the discussion, the authors speak - statistically probably correctly - of "evidence of an association", but perhaps disappointingly it is not as clear-cut as one may have hypothesized (including this reviewer's pre-reading subjective gut feeling). Perhaps there is more not easily definable confounding that couldn't be addressed in the model. Because it seems indisputable that at least the BCG / Vole bacillus UK trial had given unusually solid long-term protection against pulmonary tuberculosis among young adults and it remains baffling that stopping this scheme of adolescent vaccination should have left such little effect. Would the reasons for this phenomenon deserve a bit more extensive addressing in the discussion? That BCG vaccination has given various levels of protection (say, nothing in Chingleput) seems entirely irrelevant for the UK, here it gave a non-varying high-level of protection for about ten to 12 years before it precipitously dropped 
towards zero. This fits with what we know from effector cell immunity and it is not clear why this should not also be reflected more substantially in the period following policy change unless the models couldn't really account for all factors that matter, but it is surely not variability in vaccine protection elsewhere that is an issue here, this is only about the UK after all.

I appreciate the work of the authors with these thought-provoking study results!

Minor comments:

Several references have capitalization vs lower case issue be in names or e.g. "bcg", and also whether words in the title should be capitalized or not, etc.

References 7 and 8 are duplicates

Reference 11 is a bit of a screw-up (2018 and January-2011)

### 2019-10-07 peer-review/review-response.Rmd
added response bullets

@@
 
+*- [ ] Reviewed and corrected.*
+
 - If possible, please use square brackets for the references with the punctuation following the reference.
 
+*- [ ] Done.*
+
 - Tables should not have any empty cells as design element or because information is not available ('NA' (subsequently defined as not available or not applicable) can be used for example).
+
+*- [ ] Required*
+*- [ ] Provide link to output in form that can be coded*
 
@@
 
+*- [ ] Add*
+
 - Please check the overall text versus the tables for the name of the intervals (e.g.
@@
 To give an example: In the abstract and in the main text subsection called ‘Adjusted estimates of the effects of the change in policy on school-age children’ there is ‘IRR: 0.74 (95%CI 0.61, 0.88)’ but in the table 2 there is ‘0.74 (0.61, 0.88)’ defined as IRR (95% CrI) CrI: Credible interval.
+
+*- [ ] Check CI usage*
 
@@
 observational study?) also in the abstract (either in the aim or the methods).
+
+*- [ ] Add aim*
+*- [ ] Add study type*
+*- [ ] Cut to length*
 
@@
 
+*- [ ] Add detail on LFS*
+
 Statistical methods
@@
 
+*- [ ] Add comment on CIs in methods.*
 Ethical statement
@@
 
+*- [ ] Add ethical statement*
+
 RESULTS
@@
 = ‘Of notifications with a known birth status 27% (29096/106765) were UK born, in comparison to 33% (2634/8055) in cases with an imputed birth status.’ Just for editor’s understanding: does this mean that ‘Among notifications with a known birth status, 27% had the status ‘born in the UK’, while among notifications with an imputed birth status, 33% had the status ‘born in the UK’?
+
+*- [ ] Exlain this*
 
@@
 
+*- [ ] Say thank you*
+*- [ ] Agree with them*
+
 Introduction
@@
 1.	Paragraph 2. In France, universal vaccination was implemented in all children before entering "community care" and primary school at the latest, and not in neonates.
+
+*- [ ] Update paper with this information*
+
 Methods
@@
 1.	Paragraph 1. The geographical scale of TB local incidence is not specified and is an important piece of information that would be worth mentioning at this stage.
+
+*- [ ] Clarify what this means*
+*- [ ] Add if possible*
 
@@
 
+*- [ ] As for editor comment clarify inclusion*
+
 3.	Definition of cohorts.
@@
 Maybe further clarify this crucial point.
+
+*- [ ] Explain this example*
+*- [ ] Explain the general approach*
+*- [ ] Anything to add to the paper*
 
@@
 
+*- [ ] Thank for this and mention that it is a good point*
+*- [ ] Mention in discussion after some thought*
+
 2.	Other interventions may decrease TB incidence, such as a better screening, a change in screening policy in migrants, a better case management…This is not discussed.
 Could these potential interventions have also played a role in the results presented here?
+
+*- [ ] Agree - flag where currently mentioned*
+*- [ ] Rephrase to make clearer?*
 
@@
 A brief description of this could be given here, or in Introduction.
+
+*- [ ] Agree and discuss the low quality of the data*
+*- [ ] Explain that this motivated the form of the analysis*
 
@@
 
+*- [ ] Explain that it is the case when adults are vaccinated. If there is sufficent transmission then vaccinating adults may be more protective than vaccinating children*
+*- [ ] Rephrase to higlight this point*
+
 ## Authors' response to Reviewer #2:
@@
 
+*- [ ] Say thanks*
+
 ## Authors' response to Reviewer #3:
@@
 This study is of high relevance, not just for the England, but for the global tuberculosis community: The UK (England and Wales notably) had been one of the few countries with a universal vaccination of adolescents and a clinical trial demonstrating high efficacy for at least 20 years in the age vaccinated groups and beyond to include young adults during follow-up.
+
+*- [ ] Agree and say thank you for helpful comments*
 
@@
 This would clarify that here is a basic methodology, and here is a complex modeling description, the former for every epidemiologist, the latter only for the seasoned expert in statistical modeling using special techniques.
-I concur that it should and has to be there, but the more
-basic working techniques of calculation might be preferably "visually" separated from complex modeling techniques, notably when unexplained acronyms such as "GBM" are interspersed that do not help in easily following what the modeler actually did.
+I concur that it should and has to be there, but the more basic working techniques of calculation might be preferably "visually" separated from complex modeling techniques, notably when unexplained acronyms such as "GBM" are interspersed that do not help in easily following what the modeler actually did.
 Nor does it become that clear why exactly these models were chosen and not something else.
 The reader should probably have some inkling what the motivation for model selection is.
+
+*- [ ] Comment agreeing*
+*- [ ] Split approach into two paragraphs*
+*- [ ] As per editor correct acronyms usage and saythank you*
+*- [ ] Review explanation for model usage and update if not sufficient*
 
@@
 
+*- [ ] Agree with comment and explain approach*
+*- [ ] Review method explanation to see if can be tweaked for clarity*
+
 In table 3, among the other variables, is also the year of study eligibility (please correct spelling).
@@
 Thus perhaps the variability in the estimated magnitude of effect for cases prevented in the following paragraph which is substantial is therefore not all that meaningful after all (CrIs not uncommonly overlapping zero).
+
+*-[ ] Correct spellign of year of study eligibility - say thank you*
+*- [ ] Disagree - after adjusting for incidence rates trends would not expect a trend across years nessarily.
+Adjusting for this may account for other remaining differences between cohorts.
+Does not mean that the results are more uncertain than represented by the current CIs.*
 
@@
 
+*- [ ] Agree went against expectation of the authors*
+*- [ ] Reminder that looking at comparing two potentially effective vaccination programmes. Baseline is not no vacination*
+*- [ ] Explanation is probably low incidence rates*
+*- [ ] Also looking at a relatively small period of follow up*
+*- [ ] Major benefit is from the non-UK born whos interactons with UK vaccination is likely different*
+*- [ ] Check all points are clear in the paper*
+
 I appreciate the work of the authors with these thought-provoking study results!
@@
 
+*- [ ] Resolve*
+
 References 7 and 8 are duplicates
 
+*- [ ] Resolve*
+
 Reference 11 is a bit of a screw-up (2018 and January-2011)
+
+*- [ ] Resolve*

### 2019-10-08 peer-review/review-response.Rmd
response to initial editor comments

@@
 
-*- [ ] Reviewed and corrected.*
+*We have reviewed all abbreviations and added definitions where missing.
+Tables and footers have been checked for abbreviations without explanations and these have been added when missing.
+The abstract has had explanations for the abbreviations added where missing.*
 
@@
 
-*- [ ] Done.*
+*We have moved to using square brackets and moved references from following punctuation to punctuation following the reference.*
 
@@
 
-*- [ ] Required*
-*- [ ] Provide link to output in form that can be coded*
+*We feel that the table design we have used clearly imparts the information that it contains - part of this design is empty cells to space information.
+NA has not been used as information is not missing in the conventional sense - the results tables combine two different models with different confounders (hence not all fields contain data).
+Having this information in a single table allows for easier between model comparisons which we feel is essential to the presentation of our results.
+We have also included all results in csv form for readers who wish to use our results programmatically.
+We have clarified this in the data availability section.
+However, we are happy to alter the table design if we can preserve the clarity of the results presentation and welcome your suggestions.*
 
@@
 
-*- [ ] Add*
+*We have added location and temporal information to all studies.
+As this was a retrospective whole population based study we are not clear that sample sizes are useful in this instance (as they are the whole population meeting the cohort critiera which is age based).
+We are happy to add this if your feel this is meaningful*
 
@@
 
-*- [ ] Check CI usage*
+*We have standardised all usage across the text. *
 
@@
 
-*- [ ] Add aim*
-*- [ ] Add study type*
-*- [ ] Cut to length*
+*We have added an aim section that includes the aim of the paper and the study type. We have also made the required cuts to the abstract in order to keep within the word limit.*
 
@@
 
-*- [ ] Add comment on CIs in methods.*
+*We have added this and checked all tables and figures to make sure we are consistent.*
 Ethical statement
@@
 
-*- [ ] Exlain this*
+*Yes that is correct - we are open to rephrasing this if additional clarity is required.*
 

### 2019-10-08 peer-review/review-response.Rmd
complete to editor changes

@@
 
-*- [ ] Add detail on LFS*
+*The LFS is the largest household study in the UK.
+Whilst it is used primarily to estimate unemployment rates it is also designed to be used by local governments etc.
+to understand the structure of their population.
+It therefore provides some of the best available yearly population estimates that are stratified by ethnicity, UK birth status, and age.
+A caveat to this is that it may not capture the older adult poplulation well.
+However, this was not relevant to this study.
+We have addded these additional detials to the manuscript.*
 
@@
 *We have added this and checked all tables and figures to make sure we are consistent.*
+
 Ethical statement
@@
 
-*- [ ] Add ethical statement*
+*We have attended an ethical statement*
 

### 2019-10-08 peer-review/review-response.Rmd
completed reviewer 1 comments

@@
 
-*- [ ] Say thank you*
-*- [ ] Agree with them*
+*Thank you for your helpful review comments. We agree that understanding when/if targeted vaccination should be used rather than universal vaccination is a key TB control issue. *
 
@@
 
-*- [ ] Update paper with this information*
+*We have corrected this.*
 
@@
 
-*- [ ] Clarify what this means*
-*- [ ] Add if possible*
+*We have added that TB incidence is highly heterogeneous in England with the majority of cases occurring in urban, non-UK born, populations.*
 
@@
 
-*- [ ] As for editor comment clarify inclusion*
+*We have added additional details to this section. Please see the responses to the editors comments for an expanded discussion of the LFS.*
 
@@
 
-*- [ ] Explain this example*
-*- [ ] Explain the general approach*
-*- [ ] Anything to add to the paper*
+*We agree that this is somewhat difficult to interpret as the underlying cohort structure is complex.
+The cohorts are first split into targeted and universal vaccination based on age.
+Eligibility is then used to indicate if a given cohort was actually enrolled on a given programme.
+This gives a cohort who were enrolled and a comparison cohort who were not (resulting in a binary outcome).
+We have added additional detail outlining this to the paper.*
 
@@
 
-*- [ ] Thank for this and mention that it is a good point*
-*- [ ] Mention in discussion after some thought*
+*We agree that considering disease severity stratified by age is an important issue and could be addressed in a follow up study.
+We have added a comments detailing this to the discussion.
+We disagree on the evidence that the BCG protects mainly against severe disease in the UK where there is good evidence of strong protection against all disease with no evidence of protection weakening with increased age at vaccination.
+However, we do agree that this is relevant in other contexts.*
 
@@
 
-*- [ ] Agree - flag where currently mentioned*
-*- [ ] Rephrase to make clearer?*
+*We agree that this is a potential issue and one that we could not fully account for.
+We have partially accounted for it by adjusting for between year cohort differences but this does not account for interventions that changed at the same time as vaccination policy.
+We mentioned this limitation in the discussion,*
+
+*"Finally, we have assumed that the effect we have estimated for the change in BCG policy is due to the changes in BCG vaccination policy as well as other associated changes in TB control policy, after adjusting for hypothesised confounders.
+However, there may have been additional policy changes which we have not accounted for."*
+
+*We have now added additional details to the conclusions stressing this point further.*
 
@@
 
-*- [ ] Agree and discuss the low quality of the data*
-*- [ ] Explain that this motivated the form of the analysis*
+*We agree that vaccination coverage impacts the success/failure of vaccination programmes.
+Unfortunately, coverage estimates for the targeted vaccination programme are of very poor quality.
+This was part of the motivation for using a population based approach that looked at overall incidence rates and hence did not need to account for vaccination coverage.
+We agree that understanding coverage in targeted populations should be a focus for further work in this area.*
 
@@
 
-*- [ ] Explain that it is the case when adults are vaccinated. If there is sufficent transmission then vaccinating adults may be more protective than vaccinating children*
-*- [ ] Rephrase to higlight this point*
+*This is an interesting point and the motivator for some of our current research.
+Whilst the reviewer is correct that children rarely transmit TB this is not the case for adults.
+This means that school-age vaccination may lead to fewer overall cases in children than direct vaccination of children once onwards transmission has been accounted for due to increased protection from TB as young adults.
+This depends on the level of TB transmission in the population and therefore may not be relevant in the UK where transmission is generally thought to be low.
+As per our previous comment we disagree with the reviewer that BCG mainly impacts severe forms of TB in the UK.
+For areas where this is the case we agree that direct vaccination of children is more clearly beneficical.
+However, as effectiveness when given later in life varies so too will the thresold that school-age vaccination becomes more effective.*
 

### 2019-10-08 peer-review/review-response.Rmd
complete first draft response to reviewers

@@
 
-*- [ ] Say thanks*
+*Thank you this review.*
 
@@
 
-*- [ ] Agree and say thank you for helpful comments*
-
 From the switch from universal adolescent vaccination to selective (largely the descendant newborn of foreign-born ancestors) one would thus hypothetically expect an increase in the incidence rate among the former and a decrease in the incidence rate among the latter (among the respective cohorts in both groups), if confounding resulting from secular trends could be properly controlled for.
 This study aimed to do precisely that.
+
+*Thank you for this detailed and helpful review.
+We agree that the findings from this study should be relevant to the global TB community.
+We particularly appreciate your suggestions for improving the discussion of the approach used.*
 
@@
 
-*- [ ] Comment agreeing*
-*- [ ] Split approach into two paragraphs*
-*- [ ] As per editor correct acronyms usage and saythank you*
-*- [ ] Review explanation for model usage and update if not sufficient*
+*We agree that it is important that all readers can easily follow the analysis.
+We have split the modelling section as you suggested into a non-technical overview and a more technical discussion of the approach.
+As per the editor comment we have carefully reviewed the use of abbreviations and added definitions where missing.*spel
 
@@
 However, the authors later also state that all models (does this refer only to the non-UK born?) that included at least age in the adjustment resulted in comparable effects.
-As parsimony is king in all modeling, perhaps it could be stated more
-explicitly that always the model with the minimum required variables was used and perhaps why recourse was taken to retaining more complex models if a more simple one gave actually comparable results.
+As parsimony is king in all modeling, perhaps it could be stated more explicitly that always the model with the minimum required variables was used and perhaps why recourse was taken to retaining more complex models if a more simple one gave actually comparable results.
 Perhaps I am missing here something, but then it would indeed be useful to review the sum of the modeling exercises to the point.
 
-*- [ ] Agree with comment and explain approach*
-*- [ ] Review method explanation to see if can be tweaked for clarity*
+*We agree that parsimony is an important goal of modelling.
+However, it is also important to adjust for confounding in the data.
+Our approach used a hypothesis free (excepting our initial use of domain knowledge in selecting an overall pool of confounders), statistically rigorous metric (the leave one out information criterion) to select the model that best fit the data after being penalised for model complexity (to avoid overfitting).
+We hope that our discussion of this approach is clearer now that we have adopted your suggestion above of splitting the statistical methods.
+We included a discussion of other models as it is gives an indication of which variables greatly altered the results and may be used when interpreting the findings.
+It is also important to include to make it clear that we explored multiple models (doing otherwise might be considered p hacking).*
 
@@
 
-*-[ ] Correct spellign of year of study eligibility - say thank you*
-*- [ ] Disagree - after adjusting for incidence rates trends would not expect a trend across years nessarily.
-Adjusting for this may account for other remaining differences between cohorts.
-Does not mean that the results are more uncertain than represented by the current CIs.*
+*We feel that adjusting for year of study eligibility helps to account for potential differences between cohorts.
+We would not expect to see a trend and if we did this might indicate that we had missed adjusting for a variable that may be introducing systematic bias.
+We feel that we have satisfactorily captured uncertainty using our modelling approach and that therefore the CrIs are representive of the underlying uncertainty.
+We have corrected the spelling of eligibility.*
 
@@
 
-*- [ ] Agree went against expectation of the authors*
-*- [ ] Reminder that looking at comparing two potentially effective vaccination programmes. Baseline is not no vacination*
-*- [ ] Explanation is probably low incidence rates*
-*- [ ] Also looking at a relatively small period of follow up*
-*- [ ] Major benefit is from the non-UK born whos interactons with UK vaccination is likely different*
-*- [ ] Check all points are clear in the paper*
+*We agree that our conclusions were not as clear-cut as might be expected.
+However, part of this is due to comparing two vaccination programmes rather than using a baseline of no vaccination.
+We were in effect estimating the incremental difference between the two programmes and it is perhaps to be expected that this difference was small.
+In addition much of the uncertainty is due to incidence in neonates being very low in England throughout the study.
+It should also be noted that stopping universal school-age vaccination did result in increased incidence rates overall when only the UK born were considered (as expected).
+It required the inclusion of the non-UK born in order to alter these findings.
+The non-UK born were less clearly exposed to UK vaccination policy and as the reviewer has noted this impact may be due to other policy changes.
+We feel that the reviewer suggested changes to the discussion has helped highlight this.
+A final point is that we were only able to use 5 years of follow up.
+It may be the case that extending follow up alters these findings.
+We have included this point in the discussion.*
 
 I appreciate the work of the authors with these thought-provoking study results!
+
+*Thank you again for your useful comments.*
 
@@
 
-*- [ ] Resolve*
+*Resolved issue with BCG. Use of title case is a journal dependant decision and therefore we have not standardised.*
 
@@
 
-*- [ ] Resolve*
+*Resolved.*
 
@@
 
-*- [ ] Resolve*
+*Resolved*

### 2019-10-09 peer-review/review-response.Rmd
finalised review reponses

@@
 
-*We have added location and temporal information to all studies.
+*We have added location and temporal information to all tables/figures.
 As this was a retrospective whole population based study we are not clear that sample sizes are useful in this instance (as they are the whole population meeting the cohort critiera which is age based).
-We are happy to add this if your feel this is meaningful*
+We are happy to add this if your feel this is meaningful.*
 
@@
 However, this was not relevant to this study.
-We have addded these additional detials to the manuscript.*
+We have added these additional detials to the manuscript.*
 
@@
 
-*We have attended an ethical statement*
+*We have added an ethical statement*
 
@@
 
-*Thank you for your helpful review comments. We agree that understanding when/if targeted vaccination should be used rather than universal vaccination is a key TB control issue. *
+*Thank you for your helpful comments. We agree that understanding when/if targeted vaccination should be used rather than universal vaccination is a key TB control issue. *
 
@@
 *We agree that considering disease severity stratified by age is an important issue and could be addressed in a follow up study.
-We have added a comments detailing this to the discussion.
+We have added comments detailing this to the discussion.
 We disagree on the evidence that the BCG protects mainly against severe disease in the UK where there is good evidence of strong protection against all disease with no evidence of protection weakening with increased age at vaccination.
@@
 
-*Thank you this review.*
+*Thank you for this review.*
 
@@
 We have split the modelling section as you suggested into a non-technical overview and a more technical discussion of the approach.
-As per the editor comment we have carefully reviewed the use of abbreviations and added definitions where missing.*spel
+As per the editor comment we have carefully reviewed the use of abbreviations and added definitions where missing.*
 

### 2019-10-09 README.Rmd
updated readme

@@
 
-In 2005, England changed from universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children to targeted BCG vaccination of high-risk children at birth.
+In 2005, England changed from universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children to targeted BCG vaccination of high-risk neonates.
 
-#### Methods
+### Aim
 
-We combined notification data from the Enhanced Tuberculosis Surveillance system, with demographic data from the Labour Force Survey to construct retrospective cohorts of individuals in England relevant to both the universal, and targeted vaccination programmes between Jan 1, 2000 and Dec 31, 2010.
-For each cohort, we estimated incidence rates over a 5 year follow-up period and used Poisson and Negative Binomial regression models in order to estimate the impact of the change in policy on TB.
+Estimate the impact of the 2005 change in BCG policy on incidence rates in the targeted populations of England using an observational study.
+
+### Methods
+
+We combined notifications from the Enhanced Tuberculosis Surveillance system, with demographic data from the Labour Force Survey to construct retrospective cohorts relevant to both the universal, and targeted vaccination between Jan 1, 2000 and Dec 31, 2010.
+We then estimated incidence rates over a 5 year follow-up period and used regression modelling to estimate the impact of the change in policy on TB.
 
@@
 
-In the non-UK born, we found evidence for an association between a reduction in incidence rates and the change in BCG policy (school-age IRR: 0.74 (95%CI 0.61, 0.88), neonatal IRR: 0.62 (95%CI 0.44, 0.88)).
-We found some evidence that the change in BCG policy was associated with a increase in incidence rates in the UK born school-age population (IRR: 1.08 (95%CI 0.97, 1.19)) and weaker evidence of an association with a reduction in incidence rates in UK born neonates (IRR: 0.96 (95%CI 0.82, 1.14)).
-Overall, we found that the change in BCG policy was associated with directly preventing 385 (95% CI -105, 881) TB cases.
+In the non-UK born, we found evidence for an association between a reduction in incidence rates and the change in BCG policy (school-age incidence rate ratio (IRR): 0.74 (95% credible interval (CrI) 0.61,0.88), neonatal IRR: 0.62 (95%CrI 0.44,0.88)).
+We found some evidence that the change in policy was associated with an increase in incidence rates in the UK born school-age population (IRR: 1.08 (95%CrI 0.97,1.19)) and weaker evidence of an association with a reduction in incidence rates in UK born neonates (IRR: 0.96 (95%CrI 0.82,1.14)).
+Overall, we found that the change in policy was associated with directly preventing 385 (95%CrI -105,881) cases.
 
@@
 
-Withdrawing universal vaccination at school-age and targeting BCG vaccination towards high-risk neonates was associated with reduced incidence of TB in England.
-This was largely driven by reductions in the non-UK born.
-There was a slight increase in UK born school-age cases.
+Withdrawing universal vaccination at school-age and targeting vaccination towards high-risk neonates was associated with reduced incidence of TB.
+This was largely driven by reductions in the non-UK born with cases increasing in the UK born.
 

### 2019-11-04 peer-review/review-response.Rmd
added Ellen's comments on reviewer responses

@@
 Tables and footers have been checked for abbreviations without explanations and these have been added when missing.
-The abstract has had explanations for the abbreviations added where missing.*
+We added missing explanations for the abbreviations in the abstract*
 
@@
 
-*We have moved to using square brackets and moved references from following punctuation to punctuation following the reference.*
+*We have moved to using square brackets and moved to punctuation following the reference.*
 
@@
 
-*We feel that the table design we have used clearly imparts the information that it contains - part of this design is empty cells to space information.
-NA has not been used as information is not missing in the conventional sense - the results tables combine two different models with different confounders (hence not all fields contain data).
+*We have re-formatted the tables to limit empty cells while still clearly imparting The correct information.
+Where fields do not contain data we have indicated this with a dash (-) and only used NA for missing data in the conventional sense.
 Having this information in a single table allows for easier between model comparisons which we feel is essential to the presentation of our results.
@@
 We have clarified this in the data availability section.
-However, we are happy to alter the table design if we can preserve the clarity of the results presentation and welcome your suggestions.*
+However, we are happy to alter the table design if we can preserve the clarity of the results presentation.*
 
@@
 *We have added location and temporal information to all tables/figures.
-As this was a retrospective whole population based study we are not clear that sample sizes are useful in this instance (as they are the whole population meeting the cohort critiera which is age based).
-We are happy to add this if your feel this is meaningful.*
+As this was a retrospective whole population based study based on surveillance data, sample sizes are not useful in this instance (as they are the whole population meeting the cohort critiera which is age based).*
 
@@
 A caveat to this is that it may not capture the older adult poplulation well.
-However, this was not relevant to this study.
+However, this was not relevant to this study as those individuals were not directed impacted by the change in policy.
 We have added these additional detials to the manuscript.*
@@
 
-*Yes that is correct - we are open to rephrasing this if additional clarity is required.*
+*Yes that is correct.*
 
@@
 *We agree that this is somewhat difficult to interpret as the underlying cohort structure is complex.
-The cohorts are first split into targeted and universal vaccination based on age.
-Eligibility is then used to indicate if a given cohort was actually enrolled on a given programme.
+The cohorts are first identified based on age.
+Eligibility is then used to indicate if a given cohort was enrolled on the universal or targeted  programme.
 This gives a cohort who were enrolled and a comparison cohort who were not (resulting in a binary outcome).
@@
 1.	No comments, except that I find some parts of the results difficult to understand for a "common" public health professional" with no deep knowledge on statistics and modelling methods.
+
+*We have tried to make the results section readable but happy to take suggestions on which sections in particular to improve.*
 
@@
 *This is an interesting point and the motivator for some of our current research.
-Whilst the reviewer is correct that children rarely transmit TB this is not the case for adults.
+Whilst the reviewer is correct that children are less likely to be smear-positive this is not the case for adults.
 This means that school-age vaccination may lead to fewer overall cases in children than direct vaccination of children once onwards transmission has been accounted for due to increased protection from TB as young adults.
 This depends on the level of TB transmission in the population and therefore may not be relevant in the UK where transmission is generally thought to be low.
-As per our previous comment we disagree with the reviewer that BCG mainly impacts severe forms of TB in the UK.
-For areas where this is the case we agree that direct vaccination of children is more clearly beneficical.
-However, as effectiveness when given later in life varies so too will the thresold that school-age vaccination becomes more effective.*
+For areas where the BCG mainly impacts severe forms of TB only we agree that direct vaccination of children is more clearly beneficical.
+However, there is good evidence of effectiveness for the BCG when given later in life in the UK meaning that there is likely an incidence rate thresold at which school-age vaccination becomes more protective.*
 
@@
 *Thank you for this detailed and helpful review.
-We agree that the findings from this study should be relevant to the global TB community.
+We hope that the findings from this study should be relevant to the global TB community.
 We particularly appreciate your suggestions for improving the discussion of the approach used.*

### 2019-11-05 README.Rmd
finalised reviewer responses

@@
 
-In 2005, England changed from universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children to targeted BCG vaccination of high-risk neonates.
+In 2005 in England, universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children was replaced by targeted BCG vaccination of high-risk neonates.
 
@@
 
-Estimate the impact of the 2005 change in BCG policy on incidence rates in the targeted populations of England using an observational study.
+Estimate the impact of the 2005 change in BCG policy on tuberculosis incidence rates in England.
 
@@
 
-We combined notifications from the Enhanced Tuberculosis Surveillance system, with demographic data from the Labour Force Survey to construct retrospective cohorts relevant to both the universal, and targeted vaccination between Jan 1, 2000 and Dec 31, 2010.
+We conducted an observational study by combining notifications from the Enhanced Tuberculosis Surveillance system, with demographic data from the Labour Force Survey to construct retrospective cohorts relevant to both the universal, and targeted vaccination between Jan 1, 2000 and Dec 31, 2010.
 We then estimated incidence rates over a 5 year follow-up period and used regression modelling to estimate the impact of the change in policy on TB.

### 2019-12-13 README.Rmd
added published paper to README

@@
 [![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/DirectEffBCGPolicyChange)
-[![Paper](https://img.shields.io/badge/Paper-10.1101/567511-lightgreen.svg)](https://doi.org/10.1101/567511)
+[![Paper](https://img.shields.io/badge/Paper-10.2807%2F1560--7917.ES.2019.24.49.1900220-lightgreen)](https://doi.org/10.1016/j.vaccine.2019.06.056)
+[![Preprint](https://img.shields.io/badge/Preprint-10.1101/567511-lightgrey.svg)](https://doi.org/10.1101/567511)
 [![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)

### 2019-12-13 README.Rmd
updated link to paper:

--- 2019-11-05 README.Rmd
@@
 [![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/DirectEffBCGPolicyChange)
-[![Paper](https://img.shields.io/badge/Paper-10.1101/567511-lightgreen.svg)](https://doi.org/10.1101/567511)
+[![Paper](https://img.shields.io/badge/Paper-10.2807%2F1560--7917.ES.2019.24.49.1900220-lightgreen)](https://doi.org/10.2807/1560-7917.ES.2019.24.49.1900220)
+[![Preprint](https://img.shields.io/badge/Preprint-10.1101/567511-lightgrey.svg)](https://doi.org/10.1101/567511)
 [![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)
